Cover Times of Random Walks

### 2019-05-07 paper.tex
talk :)

\DeclareMathOperator{\E}{\mathrm{E}}		     % expected value
\DeclareMathOperator{\Var}{\mathrm{Var}}         % variance
\DeclareMathOperator{\I}{\mathbb{1}}         % variance
\DeclareMathOperator{\pr}{\mathrm{P}}		     % expected value
\DeclareMathOperator{\cov}{\uptau_\textrm{cov}}  % cover time variable
\DeclareMathOperator{\tcov}{t_\textrm{cov}}      % cover time
\DeclareMathOperator{\hit}{t_{\textrm{hit}}}     % hitting time
\DeclareMathOperator{\tmina}{t_\textrm{min}^A}   % minimum time on A
\DeclareMathOperator{\Ceff}{C_\textrm{eff}}      % effective conductance
\DeclareMathOperator{\Reff}{R_{\textrm{eff}}}    % effective resistance
\DeclareMathOperator{\Rxy}{R_{\textrm{x,y}}}     % effective resistance between x,y
\DeclareMathOperator{\T}{\mathcal{T}}            % spanning tree

@@

\abstract{
The cover time of a graph is the number of steps of a
discrete random walk until every vertex has been visited at leastonce.
This senior work provides a detailed overview
of a variety of cover time topics.
We begin by investigating the cover time of small and structured graphs but
quickly find that we need
require a less precise strategy for larger, unstructured ones.
We apply more general probablistic arguments and an important electrical
network relationship to identify bounds for arbitrary graphs.
Finally, we describe the limiting distribution of the cover time of
complete graphs with self-edges.
}

@@

\newpage
\section{Introduction and Background}

We begin by describing two intuitive examples of cover time problems.
Consider a coupon collector who opens a single box of cereal everyday.
In each box, is one of $n$ distinct coupons.
How many days until our collector has collected at least one of every coupon?
(The coupon collector's problem corresponds to the cover time of the complete
graph with self-edges.)

Next consider a similar scenario called the birthday problem.
How many people would we need to ensure that every day of the year
(excluding February 29) is represented as a birthday?
A lower bound is certainly 365 but it turns out that there is no upper bound.

@@

\subsection{Definitions}
We consider an undirected, connected graph $G$ without self- or multi-edges.
Let $V(G)$ be the set of vertices and let $E(G)$ be the set of edges on $G$.
Then $n$ is the number of vertices $|V(G)|$.
We say $i\sim j$ for $i$ and $j$ in the vertex set $V(G)$ if $(i,j)$ is in $E(G)$.
Assume that each vertex is accessible from every other vertex.
That is, there is a set of edges connecting any $x$ to any $y$.
%Without loss of generality, we label the vertices $\{1, 2, 3, ..., n\}$.

@@

			\path[->] (a) edge[loop above] (a);

@@

			\path[->] (e) edge[loop above] (e);

@@

If we let the label of each graph in \cref{fig:example}
denote the cover time from that case, we have the following system
of equations:
\begin{align}
&\tcov = 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{b}
&\textrm{a} = 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{c} \nonumber \\
&\textrm{b} = 1 + \frac{1}{2}\textrm{b} + \frac{1}{2}\textrm{e}
&\textrm{c} = 1 + 1\textrm{e} \nonumber \\
&\textrm{d} = 1 + \frac{1}{3}\textrm{a} + \frac{1}{3}\textrm{e} + \frac{1}{3}\textrm{f}
&\textrm{e} = 1 + \frac{1}{3}\textrm{c} + \frac{1}{3}\textrm{e} + \frac{1}{3}\dot 0\nonumber \\
&\textrm{f} = 1 + \frac{1}{2}\textrm{g} + \frac{1}{2}\dot 0
&\textrm{g} = 1 + \frac{2}{3}\textrm{f} + \frac{1}{3}\dot 0\nonumber \\
\end{align}

@@

We write the cover time in terms of $X_i$ and use linearity of expectation.
\begin{align}
\tcov&= \E[X_1+X_2+X_3+\cdots+X_{n+1}] = \sum_{i=1}^n \E[X_i] \nonumber
\end{align}

@@

\begin{comment}
It follows that $X_i$ is geometric with distribution
\begin{align}
\pr(X_i = k) = (1-p_i)^{k-1}p_i, \nonumber
\end{align}
where $p_i$ is the probability of moving to a previously unvisited vertex.
\end{comment}
We now find the probability $p_i$ that 
we go to a previously unvisited vertex is $(n-i)/(n-1)$.
This is because there are $n-i$ unvisited vertices and a total of $n-1$ adjacent vertices.
It follows that $\E[X_i] = (n-1)(n-i)$.
Then
\begin{align}
\tcov&=   \sum_{i=1}^{n-1} \frac{n-1}{n-i} \nonumber \\
&=  (n-1) \left(\frac{1}{n-1}+\frac{1}{n-2}+\cdots+1\right). \nonumber
\end{align}

A natural interpretation of the cover time on a complete graph is 
the so-called coupon collecting problem.
There are $r$ different coupons that are randomly packaged in cereal boxes.
An avid fan buys a box of cereal every day.
We want to know how many days until the fan has collected each coupon.
The strategy is to think of every coupon as a vertex on the complete graph.
We then move from vertex to vertex with uniform probability.
The difference between the coupon collecting problem and the complete graph
is that we have self-edges
whereas before we had to leave our current vertex at each step.
Now, we can stay in the same vertex (provided our collector found the same coupon
two days in a row).
We apply our approach to the complete graph and substitute the $r$ possible coupons
we can find for the $n-1$ vertices we could move to.
Then
\begin{align}
\tcov&=  r \left(\frac{1}{r}+\frac{1}{r-1}+\cdots+1\right). \nonumber
\end{align}

@@

Now consider the $k^{th}$ term of the sum for $1 < k \leq n-1$.
By total expectation,
\begin{align}
\E_n[T_k - T_{k-1}] &=
\sum_{i=1}^k \E_n[T_k - T_{k-1} | L(k) = \sigma_i]
P(L_k = \sigma_i). \nonumber 
\end{align}

@@

\begin{align}
\E_n[T_k - T_{k-1}] &= 0 +
\E_n[T_k - T_{k-1} | L(k) = \sigma_k]
P(L_k = \sigma_k). \nonumber 
\end{align}

@@

\section{Cover Time Distributions}\label{sec:dist}
\subsection{Complete Graph}\label{sec:complete_dist}
\cite{Du11}
We consider the distribution of the cover time of a random walk
on the complete graph with self edges.
Call $r$ the number of steps since we began the random walk.
So that the notation works out more easily,
set $r=1$ when we begin our walk at $X_0$.
Let $A_i$ be the event that vertex $i$ is not visited and
$N_n$ the number of unvisited vertices.
Then
\begin{align}
P(A_i) = (1-1/n)^r \nonumber
\end{align}
and
\begin{align}
\E[N_n] = n(1-1/n)^r. \nonumber
\end{align}

@@

We use \cref{thm:pois} to find the distribution of the cover time variable $\cov$.
Set $r = n \log n + nx$.
Then 
\begin{align}
ne^{-r/n} &= n e ^{- \log n - x} \nonumber \\
&= n e {- \log n} e^ {- x} = e^{-x}.\nonumber
\end{align}

>>>>

\DeclareMathOperator{\E}{\mathrm{E}}		     % expected value
\DeclareMathOperator{\Var}{\mathrm{Var}}         % variance
\DeclareMathOperator{\I}{\mathbb{1}}             % probability
\DeclareMathOperator{\pr}{\mathrm{P}}		     % probability 
\DeclareMathOperator{\cov}{\uptau_\textrm{cov}}  % cover time variable
\DeclareMathOperator{\tcov}{t_\textrm{cov}}      % cover time
\DeclareMathOperator{\hit}{t_{\textrm{hit}}}     % hitting time
\DeclareMathOperator{\tmina}{t_\textrm{min}^A}   % minimum time on A
\DeclareMathOperator{\Ceff}{C_\textrm{eff}}      % effective conductance
\DeclareMathOperator{\Reff}{R_{\textrm{eff}}}    % effective resistance
\DeclareMathOperator{\Rxy}{R_{\textrm{x,y}}}     % effective resistance between x,y
\DeclareMathOperator{\T}{\mathcal{T}}            % spanning tree

@@

\abstract{
The cover time of a graph is the number of steps a 
discrete random walk takes to visit vertex at least once.
We begin by investigating the cover time of small and structured graphs but
quickly find that our precise strategy fails
for larger, unstructured ones.
To bound the cover time on arbitrary graphs,
we apply more general probablistic arguments and an important electrical
network identity.
Finally, we describe the limiting distribution of the cover time of
complete graphs with self-edges and apply our analysis to an intuitive example.
}

@@

\newpage
\section{Introduction and Background}\label{sec:intro}

To introduce the concept of a cover time, we begin by describing two intuitive examples.
Consider a coupon collector who receives one of $r$ distinct coupons
in the mail everyday.
For example if $r=3$, the collector may receive coupon 1 today, coupon 3 tomorrow,
coupon 1 (again) the next day, and finally coupon 2 the following day.
For arbirtrary $r$, we ask how many days until our 
collector has collected at least one of every coupon?
(To solve the coupon collector problem we will frame it in terms 
of the complete graph with self-edges.)

Next consider a particular case of coupon collecting called the birthday problem.
How many people would we need to ensure that every day of the year
(excluding February 29) is represented as a birthday?
A lower bound is certainly 365 but there is no upper bound.

@@

\subsection{Definitions}
We consider an undirected, connected graph $G$ without multi-edges.
Let $V(G)$ be the set of vertices and let $E(G)$ be the set of edges on $G$.
Then $n$ is the number of vertices $|V(G)|$.
We say $i\sim j$ for $i$ and $j$ in the vertex set $V(G)$ if $(i,j)$ is in $E(G)$.
Assume that each vertex is accessible from every other vertex.
That is, there is a set of edges connecting any $x$ to any $y$.

@@

			\path[->] (a) edge[loop above, distance=6mm] (a);

@@

			\path[->] (e) edge[loop above, distance=6mm] (e);

@@

If we let the label of each graph in \cref{fig:example}
denote the cover time from that case, we have the following system
of equations:
\begin{align}
&\tcov = 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{b}
&\textrm{a} = 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{c} \nonumber \\
&\textrm{b} = 1 + \frac{1}{2}\textrm{b} + \frac{1}{2}\textrm{e}
&\textrm{c} = 1 + 1\textrm{e} \nonumber \\
&\textrm{d} = 1 + \frac{1}{3}\textrm{a} + \frac{1}{3}\textrm{e} + \frac{1}{3}\textrm{f}
&\textrm{e} = 1 + \frac{1}{3}\textrm{c} + \frac{1}{3}\textrm{e} + \frac{1}{3}\dot 0\nonumber \\
&\textrm{f} = 1 + \frac{1}{2}\textrm{g} + \frac{1}{2}\dot 0
&\textrm{g} = 1 + \frac{2}{3}\textrm{f} + \frac{1}{3}\dot 0\nonumber
\end{align}

@@

We write the cover time in terms of $X_i$ and use linearity of expectation.
\begin{align}
\tcov&= \E[X_1+X_2+X_3+\cdots+X_{n-1}] = \sum_{i=1}^{n-1} \E[X_i] \nonumber
\end{align}

@@

\begin{comment}
It follows that $X_i$ is geometric with distribution
\begin{align}
\pr(X_i = k) = (1-p_i)^{k-1}p_i, \nonumber
\end{align}
where $p_i$ is the probability of moving to a previously unvisited vertex.
\end{comment}
We now find the probability $p_i$ that 
we go to a previously unvisited vertex is $(n-i)/(n-1)$.
This is because there are $n-i$ unvisited vertices and a total of $n-1$ adjacent vertices.
It follows that $\E[X_i] = (n-1)/(n-i)$.
Then
\begin{align}
\tcov&=   \sum_{i=1}^{n-1} \frac{n-1}{n-i} \nonumber \\
&=  (n-1) \left(\frac{1}{n-1}+\frac{1}{n-2}+\cdots+1\right). \nonumber
\end{align}

A natural interpretation of the cover time on a complete graph is 
the coupon collector problem we introduced in \cref{sec:intro}.
There are $r$ distinct coupons one of which arrives in the mail everyday.
We want to know how many days until the collector has collected each coupon.
The strategy is to think of every coupon as a vertex on the complete graph.
We then move from vertex to vertex with uniform probability.
The difference between the coupon collector problem and the complete graph
is that we now have self-edges
whereas before we had to leave our current vertex at each step.
That is, we can stay in the same vertex (if the collector receives the same coupon
two days in a row).
We apply our approach to the complete graph and substitute the $r$ possible coupons
we can find for the $n-1$ vertices we could move to.
Then
\begin{align}
\tcov&=  r \left(\frac{1}{r}+\frac{1}{r-1}+\cdots+1\right). \nonumber
\end{align}

@@

Now consider the $k^{th}$ term of the sum for $1 < k \leq n-1$.
By total expectation,
\begin{align}
\E_n[T_k - T_{k-1}] &=
\sum_{i=1}^k \E_n[T_k - T_{k-1} | L(k) = \sigma_i]
\pr(L_k = \sigma_i). \nonumber 
\end{align}

@@

\begin{align}
\E_n[T_k - T_{k-1}] &= 0 +
\E_n[T_k - T_{k-1} | L(k) = \sigma_k]
\pr(L_k = \sigma_k). \nonumber 
\end{align}

@@

\section{Cover Time Distributions}\label{sec:dist}
\subsection{Complete Graph}\label{sec:complete_dist}
\cite{Du11}
We consider the distribution of the cover time of a random walk
on the complete graph with self edges.
Call $r$ the number of steps since we began the random walk.
So that the notation works out more easily,
set $r=1$ when we begin our walk at $X_0$.
Let $A_i$ be the event that vertex $i$ is not visited and
$N_n$ the number of unvisited vertices.
Then
\begin{align}
\pr(A_i) = (1-1/n)^r \nonumber
\end{align}
and
\begin{align}
\E[N_n] = n(1-1/n)^r. \nonumber
\end{align}

@@

We use \cref{thm:pois} to find the distribution of the cover time variable $\cov$.
Set $r = n \log n + nx$.
Then 
\begin{align}
ne^{-r/n} &= n e ^{- \log n - x} \nonumber \\
&= n e^{- \log n} e^ {- x} = e^{-x}.\nonumber
\end{align}
